Supporting information: Tail associations in ecological variables and their impact on extinction risk

### 2020-01-25 SuppMat.Rmd
worked on comments and edits
table cleaned
typo

---
title: "Supporting information: Tail associations and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

dt2<-tibble("$\\lambda_i(t)$" = c("$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
                          "$\\exp(\\epsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
))

@@

**The Hassell model.** The Hassell model [@hassell1975density; @may1976bifurcations] was 
developed to help understand the dynamics of insect populations, and as a model of contest versus scramble 
instraspecific competition. The model
can be written $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$. Here $r>0$ is the intrinsic growth rate and
$a$ and $b$ are positive parameters. If $b=1$, the Hassell model reduces to the
Beverton-Holt model [@beverton2012dynamics] used in fisheries. The Hassell model is a representation
of contest competition if $b \sim 1$, and scramble competition if $b$ is large.
Like the Ricker model, the Hassell model also has two equilibria, $P_e =0$ 
(the trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
For $P_e = P^\star$, the model shows undercompensatory dymamics if $1<r<[b/(b-1)]^b$ 
and overcompensatory dynamics if $[b/(b-1)]^b<r<[b/(b-2)]^b$, so $r_{min}=1$, 
$r_c = [b/(b-1)]^b$, and $r_{bf}=[b/(b-2)]^b$.

>>>>

---
title: "Supporting information: Tail associations in ecological variables and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

dt2<-tibble("$\\lambda_i(t)$" = c("$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} \\exp[r (1-P_i(t)/K)$]",
                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/(1+aP_i(t))^b$",
                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/[1+(aP_i(t))^b]$",
                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} r/[1+\\exp \\{-a(1-P_i(t)/b)\\}]$",
                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} [1+r(1-P_i(t)/K)]$",
                          "$\\exp(\\varepsilon_i(t)) \\hspace{0.1 cm} [1+r(K-P_i(t))\\{1-\\exp(-sP_i(t))\\}]$"
))

@@

**The Hassell model.** The Hassell model [@hassell1975density; @may1976bifurcations] was 
developed to help understand the dynamics of insect populations, and as a model of contest versus scramble 
intraspecific competition. The model
can be written $P(t+1) = F(P(t),r,a,b)=rP(t)/(1+aP(t))^b$. Here $r>0$ is the intrinsic growth rate and
$a$ and $b$ are positive parameters. If $b=1$, the Hassell model reduces to the
Beverton-Holt model [@beverton2012dynamics] used in fisheries. The Hassell model is a representation
of contest competition if $b \sim 1$, and scramble competition if $b$ is large.
Like the Ricker model, the Hassell model also has two equilibria, $P_e =0$ 
(the trivial extinction equilibrium) and $P_e = P^\star =(r^{1/b}-1)/a$. 
For $P_e = P^\star$, the model shows undercompensatory dymamics if $1<r<[b/(b-1)]^b$ 
and overcompensatory dynamics if $[b/(b-1)]^b<r<[b/(b-2)]^b$, so $r_{min}=1$, 
$r_c = [b/(b-1)]^b$, and $r_{bf}=[b/(b-2)]^b$.
